---
title: "Reproducible Research: Peer Assessment 1"
output: "PA1_template_files/figure-html"
  html_document:
    keep_md: true
---
## load libiary 
dplyr,data.table

```{r, echo=FALSE}

library(dplyr)
library(data.table)

```


## Loading and preprocessing the data (Remove NAs)

```{r unzipfile, echo=TRUE}

unzip(zipfile = "activity.zip", exdir = getwd())

dt <- read.table("activity.csv", na.strings = "NA", sep = "," , header = T)

dt <- dt[complete.cases(dt),]

dt$date <- as.Date(dt$date, format = "%Y-%m-%d")


```


## What is mean total number of steps taken per day?

### Aggreated date by Days

```{r allstepsperday ,echo=TRUE}
totalStepbyDay <- aggregate(dt$steps, by=list(date = dt$date), FUN = sum)

colnames(totalStepbyDay)[2] <- "daily_steps"

# print summary of daily steps
summary(totalStepbyDay$daily_steps)

```
### Draw the histgram, Median and Mean lines (Median and Mean are overlapping)

```{r, echo=TRUE}
library(ggplot2)
# Basic histogram
myplot <- ggplot(totalStepbyDay, aes(daily_steps)) + geom_histogram(binwidth=2000)

myplot + 
  
  geom_vline(xintercept = mean(totalStepbyDay$daily_steps),        # Add line for mean
             col = "red",
             lwd = 3) + 
  
  annotate("text",                        # Add text for mean
           x = 16000,
           y = 8,
           label = paste("Mean =", mean(totalStepbyDay$daily_steps)),
           col = "red",
           size = 5) + 
  
  geom_vline(xintercept = median(totalStepbyDay$daily_steps),        # Add line for mean
             col = "blue",
             lwd = 1) +


    annotate("text",                        # Add text for mean
           x = 14000,
           y = 7,
           label = paste("Median  =", median(totalStepbyDay$daily_steps),"(Media and Mean overlaps)"),
           col = "blue",
           size = 5)
  

```


## What is the average daily activity pattern?

### aggregate data by 5min interval 

```{r, echo=TRUE}
dailyStepInterval <- aggregate(dt$steps, by=list(dailyInterval = dt$interval), FUN = sum)

colnames(dailyStepInterval)[2] <- "daily_steps"

```

### Plot interval and total steps

```{r, echo=TRUE}
ggplot(data=dailyStepInterval, 
       aes(x=dailyInterval, 
           y=daily_steps, group=1)) +
            labs(x = "5 min interval")+
  
            geom_line()+
            geom_point()

```



## Imputing missing values

### Checking NA in orignal data set

```{r, echo=TRUE}

#read in original dt
dirtydt <- read.table("activity.csv", na.strings = "NA", sep = "," , header = T)

summary(dirtydt)

```


### Replace NA with interval-median (2304 NA in "setpes" accroding to summary)

```{r, echo=TRUE}

# Create a table with interval and medial of steps 
interMedian <- aggregate(dt$steps, by=list(interval = dt$interval), FUN = median)

colnames(interMedian)[2] <- "steps"

# keep the dirty data, just for comparision
tempCol <- dirtydt$steps
dirtydt$OrgStep <- tempCol

# left join dirty table with interMedian on interval
# replace steps in dirtydt with NA with steps in interMedian table
# remove -steps.y, -steps.x 
temptbl <- left_join(dirtydt,interMedian, by = "interval") %>% 
        mutate(steps = ifelse(is.na(steps.x), steps.y, steps.x)) %>%
        select(-steps.y, -steps.x) 


totalStepbyDay1 <- aggregate(temptbl$steps, by=list(date = temptbl$date), FUN = sum)

colnames(totalStepbyDay1)[2] <- "daily_steps"


str(dirtydt)

```


### plot for imputed NA data -- Mean, median are both different
```{r, echo=TRUE}

myplot1 <- ggplot(totalStepbyDay1, aes(daily_steps)) + geom_histogram(binwidth=2000)

myplot1 + 
  
  geom_vline(xintercept = mean(totalStepbyDay1$daily_steps),        # Add line for mean
             col = "red",
             lwd = 1) + 
  
  annotate("text",                        # Add text for mean
           x = 16000,
           y = 8,
           label = paste("Mean =", mean(totalStepbyDay1$daily_steps)),
           col = "red",
           size = 5) + 
  
  geom_vline(xintercept = median(totalStepbyDay1$daily_steps),        # Add line for mean
             col = "blue",
             lwd = 1) +


    annotate("text",                        # Add text for mean
           x = 14000,
           y = 7,
           label = paste("Median  =", median(totalStepbyDay1$daily_steps),"(Media and Mean overlaps)"),
           col = "blue",
           size = 5)
  

```



## Are there differences in activity patterns between weekdays and weekends?

### adding factor var

```{r, echo=TRUE}

# fix date format
temptbl$date <- as.Date(temptbl$date, format = "%Y-%m-%d")


# convert date to weekeday, change to factor var
temptbl$Weekdays <- weekdays(temptbl$date)

temptbl <- temptbl %>% mutate(Weekdays = ifelse(Weekdays %in% c("Saturday","Sunday"), "Weekend", "Weekday")
                              ,as.factor(Weekdays))


# Select sum of steps, group by interval and weekedays 
q5data <- temptbl %>% 
                  group_by(interval, `as.factor(Weekdays)`) %>%
                  summarise(totaldailysteps = sum(steps))

```
### produce the plot, over lay weekday and weekend data

```{r, echo=TRUE}

ggplot(q5data, aes(x=interval,y=totaldailysteps,group=`as.factor(Weekdays)`)) + 
      geom_line(aes(color=`as.factor(Weekdays)`))

```

